---
title: "Practical Machine Learning Assignment"
author: "finbark"
date: "30 September 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
```

# Overview
This report outlines predictions for the manner in which six participants performed a particular type of exercise as based on data from a fitness tracker. 20 differnet test cases are predicted using the prediction model.

# Getting the data

```{r get}
set.seed(12345)
trainUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
predictionUrl <- "http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

# Load and tidy data
Load the data and standardise how the data represents NA values.
```{r load}
trainingData <- read.csv(url(trainUrl), na.strings = c("NA", "#DIV/0!", ""))
predictionData <- read.csv(url(predictionUrl), na.strings = c("NA", "#DIV/0!", ""))
```

Remove variables one to six as they are not relevant.
```{r remove}
trainingData <- trainingData[, -(1:6)]
predictionData <- predictionData[, -(1:6)]
```

Remove variables that are more than 70% NAs.
```{r tidy}
NAs <- sapply(trainingData, function(x) mean(is.na(x))) > 0.7
trainingData <- trainingData[, NAs == FALSE]
predictionData <- predictionData[, NAs == FALSE]
```

Partition into training and testing sets.
```{r partition}
inTraining <- createDataPartition(trainingData$classe, p = 0.6, list = FALSE)
trainingSet <- trainingData[inTraining, ]
testSet <- trainingData[-inTraining, ]
```

# Predicting with random forests

We shall first try to predict with random forests as they are know to be quite effective. 
```{r randomForest}
modelFit <- randomForest(classe ~ ., data = trainingSet)
prediction <- predict(modelFit, testSet, type = "class")
matrix <- confusionMatrix(prediction, testSet$classe)
matrix
plot(modelFit)
```

The above matrix and plot show that the random forest does a good job at predicting the classe variable as it has a high accuracy score of 0.9975. 

# Predicting class on test data
``` {r predictTest}
prediction <- predict(modelFit, predictionData, type = "class")
prediction
```
